---
title: "Facets Examples"
output: html_document
---
Facets allows the user to see different types of summaries. For example, facets of a string fields, such as "gene_symbol" returns the 10 most common gene_symbols in the dataset and their counts. Facets, may be used to return percentiles, histograms and statistics depending on the queried field. These are a few examples of how facets can be used in SolveBio.

```{r}
library(solvebio)
library(data.table)

# log into your SolveBio account
login()

# get the top 10 most common gene_symbols and clinical significance for the ClinVar/Combined dataset
ClinVar <- Dataset.get_by_full_path("solveBio:public:/ClinVar/3.7.4-2017-01-30/Combined-GRCh37")
facets <- Dataset.facets(id = ClinVar$id, facets = list("clinical_significance", "gene_symbol"))

# convert the outputs into corresponding data tables:
clin_sign_table <- as.data.table(facets$clinical_significance)
gene_symbol_table <- as.data.table(facets$gene_symbol)

# assign the names to the columns of data tables
names(clin_sign_table) <- c("clinical_significace","count")
names(gene_symbol_table) <- c("gene_symbol","count")

# counts are returned as strings, so convert to integers for further analysis, eg
barplot(as.integer(gene_symbol_table$count),
        main = "Top 10 occurring Gene Symbol counts",
        names.arg = gene_symbol_table$gene_symbol,
        cex.names=0.6)

mean_count <-mean(as.integer(gene_symbol_table$count))

```
---
Using histogram and percentile facets.
```{r}
# return quantile information for genomic_coordinates.start field:
perc <- Dataset.facets(ClinVar$id, '{"genomic_coordinates.start": {"facet_type": "percentiles"}}')

#return a histogram for genomic_coordinates.start field
hist <- Dataset.facets(ClinVar$id, 
                      '{"genomic_coordinates.start": 
                      {"facet_type": "histogram"}}')

#One can use facets to specify the interval of the histogram, for example by calculating the Freedman–Diaconis (FD) binsize:
# first, get the number of total records
total_doc <- ClinVar$documents_count

# calculate the interquantile range:
iqr <- perc$genomic_coordinates.start$`75.0` - perc$genomic_coordinates.start$`25.0`

# use the FD formula to get the binsize
binsize <- 2*iqr/(total_doc^(1/3)) # binsize = 3112312

# set the interval equal to the binsize
hist <- Dataset.facets(ClinVar$id, 
                      '{"genomic_coordinates.start": 
                      {"facet_type": "histogram", "interval":"3112312"}}')

#plot the facets histogram:
plot(hist$genomic_coordinates.start[,1],
     hist$genomic_coordinates.start[,2],
     type = "h",
     main = "genomic_coordinates.start: number of occurences",
     ylab = "count",
     xlab = "")

```

Alternately, R has a built-in function to calculate the intervals, but in that case one needs to download the query first, which may take a little longer. To speed up the download, only the field of interest can be queried.
```{r}
# load the whole dataset
whole_query <- Dataset.query(ClinVar, paginate=TRUE, limit = 10000)

# alternately, load only the field of interest, e.g. genomic_coordinates.start to make it faster:
gen_coord_start_query <- Dataset.query(ClinVar, paginate=TRUE, limit = 10000,
                                      fields=list("genomic_coordinates.start"))

# R can calculate the Freedman–Diaconis binsize = use breaks = "FD"
hist <- hist(whole_query$genomic_coordinates.start,
            breaks="FD",
            main = "Genomic.coordinates.start: count of term occurences",
            ylab = "count",
            xlab = "")
```
