---
title: "Global Beacon and Global Search"
output:
  html_document:
    toc: true
    toc_depth: 4
    df_print: paged
---

## Global Beacon

Global Beacon lets anyone in your organization find datasets based on the entities it contains (i.e. variants, genets, targets). Only datasets that contain entities can be indexed.

```{r}
# Login to SolveBio via API key
library("solvebio")
#login(api_key="YOUR_API_KEY")
```

First let's start with enabling Global Beacon on the dataset:
```{r}
# Dataset ID
dataset_id <- "1658666726768179211"

# Turn on Global Beacon on the selected dataset
Dataset.enable_global_beacon(dataset_id)
```

Let's check now the status of Global Beacon indexing for the datasets:
```{r}
# Waiting a minute until indexing is complete
Sys.sleep(60)

# Getting the status of global beacon on the dataset
Dataset.get_global_beacon_status(dataset_id)
```
As we can see from the response, the indexing in complete. It means that now we can perform search on dataset entities.

Similarly to `enabling` Global Beacon for the datasets, you may also `disable` it using `disable_global_beacon` function:

```{r}
# Dataset.disable_global_beacon(dataset_id)
```

## Global Search

Global Search allows you to search for vaults, files, folders, and datasets by name, tags, user, date, and other metadata which can be customized.
Global Search is available on SolveBio Mesh. For more information about Global Search please look at [SolveBio docs](https://docs.solvebio.com/search/).

Similarly to Global Search on the web application, the search functionality is available through solvebio R client as well.

### Global Search functions

`GlobalSearch` module provides three functions and all of them use the same set of parameters (filters, entities, query, limit, offset, etc):

#### 1. GlobalSearch.search
   Performs a global search based on provided filters, entities, queries (advanced query) and returns an R data frame containing results from API response. For full list of results set `paginate` attribute to TRUE:

   ```R
   GlobalSearch.search(filters = '[{"and":[["type__in",["dataset"]]]}]')
   GlobalSearch.search(query = "fuji")
   GlobalSearch.search(query = "fuji", paginate=TRUE)
   ```

#### 2. GlobalSearch.subjects
   Similar to the search function, returns the results in form of R data frame. Returned data frame contains subjects:

   ```R
   GlobalSearch.subjects(entities = '[["gene","BRCA2"]]')
   ```
   
#### 3. GlobalSearch.request
   Performs low-level global search based on the provided filters, similar as you would do from web application.
   In the response there will be the following attributes:
   
   - results - list of vault objects (datasets, files, folders and vaults). Those are the same objects from the "Results" tab in GlobalSearch page on Mesh.
   - total - number of objects in the search results
   - vaults - list of vaults
   - subjects - list of subjects
   - subjects_count - number of subjects in the subjects list
   - took - amount of time it took to perform search
   - offset - offset for pagination, offset from the first result you want to fetch
   
   You may call the request function by providing some of the following arguments: filters, entities, query (advanced search query), limit, offset:

   ```R
   GlobalSearch.request(query = "fuji", limit=200)
   GlobalSearch.request(entities = '[["gene","BRCA2"]]')
   GlobalSearch.request(entities = '[["gene","BRCA2"]]', filters = '[{"and":[{"and":[["created_at__range",["2021-11-28","2021-12-28"]]]},["type__in",["dataset"]]]}]')
   ```


`Note`: Recomended functions to use are:

- GlobalSearch.search - for getting the search results.
- GlobalSearch.subjects - for getting the subjects.

### Examples


#### 1. Global Beacon Search

As we previously indexed the dataset, we should be able to perform entity search and see that dataset in the results.
```{r}
results <- GlobalSearch.search(entities = '[["gene","BRCA2"]]')
results
```

Each result object has the following attributes:
```{r}
names(results)
```


#### 2. Applying filters for Global Search

Search only for vaults:
```{r}
response <- GlobalSearch.search(filters = '[{"and":[["type__in",["vault"]]]}]')
response
```

Search based on date created:
```{r}
response <- GlobalSearch.search(filters = '[{"and":[{"and":[["created_at__range",["2021-11-21","2021-12-28"]]]}]}]')
response
```


#### 3. Advanced search query

The `search` function has embedded pagination in itself, so by setting that attribute to true, it will fetch all results (warning: that operation may be costly and time consuming).

By performing advanced search using `query` argument, only 100 objects are returned and you may see the reason for that in the following message:
```{r}
response <- GlobalSearch.search(query = "fuji")
dim(response)
```

We can perform a request function call to get the full API response and see how many results we have in total:
```{r}
response <- GlobalSearch.request(query = "fuji")
response$total
```

There are 1407 object in total but we have only 100 of them returned as default. In order to get all the results you may use parameter `paginate = TRUE` (please note that retrieving all objects may take a while):

```{r}
results <- GlobalSearch.search(query = "fuji", paginate = TRUE)
print(dim(results))
```


Alternatively, instead of using `paginate` parameter you may use the `limit` parameter instead. Here we're setting limit to 500 objects:
```{r}
results <- GlobalSearch.search(query = "fuji", limit=500)
print(dim(results))
```


#### 4. Getting the Global Search subjects

Similar as in the `search` function to get the result objects in the previous sections, we can use `subjects` function to get a data frame containing only subjects:

```{r}
GlobalSearch.subjects(entities = '[["gene","BRCA2"]]')
```

