---
title: "Global Beacon and Global Search"
output:
  html_document:
    toc: true
    df_print: paged
---

## Global Beacon

Global Beacon lets anyone in your organization find datasets based on the entities it contains (i.e. variants, genets, targets). Only datasets that contain entities can be indexed.

```{r}
# Login to SolveBio via API key
library("solvebio")
#login(api_key="YOUR_API_KEY")
```

First let's start with enabling Global Beacon on the dataset:
```{r}
# Dataset ID
dataset_id <- "1658666726768179211"

# Turn on Global Beacon on the selected dataset
Object.enable_global_beacon(dataset_id)
```

Let's check now the status of Glboal Beacon indexing for the datasets:
```{r}
# Waiting a minute until indexing is complete
Sys.sleep(60)

# Getting the status of global beacon on the dataset
Object.get_global_beacon_status(dataset_id)
```
As we can see from the response, the indexing in complete. It means that now we can perform search on dataset entities.

Similarly to `enabling` Global Beacon for the datasets, you may also `disable` it using `disable_global_beacon` function:
```R
Object.disable_global_beacon("1676139881237207342")
``` 


## Global Search

Global Search allows you to search for vaults, files, folders, and datasets by name, tags, user, date, and other metadata which can be customized.
Global Search is available on SolveBio Mesh. For more information about Global Search please look at [SolveBio docs](https://docs.solvebio.com/search/).

Similarly to Global Search on the web application, the search functionality is available through solvebio R client as well.

### Global Search functions

`GlobalSearch` module provides three functions:

1. `GlobalSearch.search` - Performs global search based on the provided filters, similar as you would do from web application.
   In the response there will be the following attributes:
   
   - results - list of vault objects (datasets, files, folders and vaults). Those are the same objects from the "Results" tab in GlobalSearch page on Mesh.
   - total - number of objects in the search results
   - vaults - list of vaults
   - subjects - list of subjects
   - subjects_count - number of subjects in the subjects list
   - took - amount of time it took to perform search
   - offset - offset for pagination, offset from the first result you want to fetch
   
   You may call the seach function by providing some of the following arguments: filters, entities, query (advanced search query), limit, offset:

   ```R
   GlobalSearch.search(query = "fuji", limit=200)
   GlobalSearch.search(entities = '[["gene","BRCA2"]]')
   GlobalSearch.search(entities = '[["gene","BRCA2"]]', filters = '[{"and":[{"and":[["created_at__range",["2021-11-28","2021-12-28"]]]},["type__in",["dataset"]]]}]')
   ```

2. `GlobalSearch.results` - Similar as search function, performs a global search based on provided filters, entities, queries but returns an R data frame containing results from API response. For full list of results set `paginate` attribute to TRUE:

   ```R
   GlobalSearch.results(filters = '[{"and":[["type__in",["dataset"]]]}]')
   GlobalSearch.results(query = "fuji")
   GlobalSearch.results(query = "fuji", paginate=TRUE)
   ```

3. `GlobalSearch.subjects` - Like results function returns results in form of R data frame, subjects function returns R data frame containing subjects. In order to get subjects from search you must provide `entities` search parameter:

   ```R
   GlobalSearch.subjects(entities = '[["gene","BRCA2"]]')
   ```

### Examples

#### 1. Global Beacon Search

As we previously indexed the dataset, we should be able to perform entity search and see that dataset in the results.
```{r}
response <- GlobalSearch.search(entities = '[["gene","BRCA2"]]')
```

Response attributes:
```{r}
names(response)
```

Response results (we can see that in the results there is the dataset which was previously indexed):
```{r}
response$results
```

Response subjects:
```{r}
response$subjects
```

Full response structure:
```{r}
response
```


From the response, we can see that we have 3 results in total (`total` field). Results themselves could be accessed through `results` field in the response.
The total subject count for this search is one (`subjects_count`).

#### 2. Applying filters for Global Search

Search only for vaults:
```{r}
response <- GlobalSearch.search(filters = '[{"and":[["type__in",["vault"]]]}]')
response$results
```

Search based on date created:
```{r}
response <- GlobalSearch.search(filters = '[{"and":[{"and":[["created_at__range",["2021-11-21","2021-12-28"]]]}]}]')
response$results
```


#### 3. Advanced search query:

By performing advanced search as we can see, we have 1407 objects in total but only 100 are returned:
```{r}
response <- GlobalSearch.search(query = "fuji")
print(paste('Objects in total: ', response$total))
print(paste('Results: ', dim(response$results)[1]))
```

We can either set `limit` attribute to fetch them all:
```{r}
response <- GlobalSearch.search(query = "fuji", limit=2000)
print(paste('Results: ', dim(response$results)[1]))
```

Or we can use `results` function with argument `parinate = TRUE`:
```{r}
results <- GlobalSearch.results(query = "fuji", paginate = TRUE)
print(paste('Results: ', dim(results)))
```

```{r}
results
```


#### 4. Getting the Global Search results

We have already seen in previous section how we can use `results` function. That function has embeded pagination in itself, so by setting that attribute to true, it will fetch all results (warning: that operation may be costly and time consuming).

Directly get only results from search:
```{r}
GlobalSearch.results(filters = '[{"and":[["type__in",["dataset"]]]}]')
```

Not all results are returned:
```{r}
results <- GlobalSearch.results(query = "fuji")
dim(results)
```

Adding `paginate=TRUE` (this call takes a bit more time to fetch all of the results):
```{r}
results <- GlobalSearch.results(query = "fuji", paginate=TRUE)
dim(results)
```


#### 5. Getting the Global Search subjects

Similar as in the results section above, we can use `subjects` function to get a data frame containing only subjects:

```{r}
GlobalSearch.subjects(entities = '[["gene","BRCA2"]]')
```

