---
title: "SolveBio R Tutorial"
output: html_document
---
Install the R solvebio package:
---
https://github.com/solvebio/solvebio-r#installation
---
install.packages('solvebio')
---
Load the solvebio package:
```{r,message = FALSE}
library(solvebio)
```
We'll use the ICGC Donor dataset. You can explore this dataset in your browser with https://my.solvebio.com/data/ICGC/2.0.0-21/Donor.
Retrieve and access data about a dataset:
```{r, message = FALSE}
icgc_donor = Dataset.query('ICGC/2.0.0-21/Donor', paginate=TRUE,limit = 10000)
```
We'll set the initial Kaplan-Meier interval sizes and total interval sizes as well as our initial filters. This particular filter will compare survival curves between the total ICGC dataset (every patient with survival information) and a subset of the ICGC that begins with the project code PACA (pancreatic cancer projects).
```{r,message = FALSE}
# interval sizes are in days
interval_size = 90
total_interval_to_follow = 1825
#filter by donor survival time for the control group
#filter1 = '[["donor_survival_time__gt",0]]' 
filter1 = '[{"not":["project_code__prefix","PACA"]},["donor_survival_time__gt",0]]'
#query1 = Dataset.query("ICGC/2.0.0-21/Donor", filters=filters1,paginate=TRUE,limit = 10000)

#filter for the survival time and and project code to get the the query for PACA
filter2 = '[["project_code__prefix","PACA"],["donor_survival_time__gt",0]]'
#run the initial queries for both datasets
icgc_donor1 = Dataset.query('ICGC/2.0.0-21/Donor', paginate=TRUE,limit = 10000,filters = filter1) #control dataset
icgc_donor2 = Dataset.query('ICGC/2.0.0-21/Donor', paginate=TRUE,limit = 10000,filters = filter2) #PACA dataset
#get the total record count for both datasets
count1 = dim(icgc_donor1)[1]
count2 = dim(icgc_donor2)[1]
```
Now we create the vectors for the PACA and control datasets and query ICGC for the survival numbers at each interval.
```{r,message = FALSE}
#create the x and y vectors to be plotted for both datasets
# x will be for time, which starts at 0 (as the initial filter)
# y will be for percent survival, which starts at 100%
f1_x = c(0)
f1_y = c(100)
f2_x = c(0)
f2_y = c(100)
#fill in the vectors with values 
for(day in seq(interval_size,total_interval_to_follow,interval_size)) {
  #control dataset
  f1 = list(list("donor_survival_time__gt",day)) #filter the query depending on the interval time given (day)
  q1 = Dataset.query('ICGC/2.0.0-21/Donor', paginate=TRUE,limit = 10000,filters = f1) #control dataset
  c1 = dim(q1)[1]
  #get the % total survival at current interval
  f1_percent_alive = (100*c1/count1)
  # populate the control x and y vectors with values for the current interval
  f1_x  <- c(f1_x, day)
  f1_y  <- c(f1_y, f1_percent_alive)
  # same PACA dataset
  f2 = list(list("project_code__prefix","PACA"),list("donor_survival_time__gt",day)) # add filter for PACA
  q2 = Dataset.query('ICGC/2.0.0-21/Donor', paginate=TRUE,limit = 10000,filters = f2)
  c2 = dim(q2)[1]
  f2_percent_alive = (100*c2/count2)
  f2_x  <- c(f2_x, day)
  f2_y  <- c(f2_y, f2_percent_alive)
}
```
Now we plot the survival data for each group in a Kaplan-Meier plot. This graph shows the drop-off in survival rates for the two groups, control and PACA.
```{r,message = FALSE}
plot(f1_x,f1_y,t="S",col="blue",main = "Kaplan-Meier Survival Curve",xlab="Time",ylab="Percent survival",xaxt='n')
axis(side=1, at=seq(0, total_interval_to_follow, by=interval_size))
lines(f2_x,f2_y,t="S",col="orange")
legend(1500,99,legend=c("Control", "PACA"),col=c("blue","orange"),bty='n', cex=.8,lty=1)
```